Prediction of 2024 US election ...

### 2024-11-03 paper/paper.qmd
fixed electoral college map legend

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  group_by(state, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "State",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  geom_vline(aes(xintercept = 50), color = purple, size = 1) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

@@

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  filter(!is.na(methodology)) %>% 
  group_by(methodology, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = methodology, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "Methodology",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  geom_vline(aes(xintercept = 50), color = purple, size = 1) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

```
Figure @fig-avg-pct-methodology indicates minimal bias in polling percentages when comparing different polling methodologies, as most polls show an even distribution of support between candidates. However, polls that incorporate IVR (Interactive Voice Response) and online panels tend to exhibit greater variability, often resulting in more pronounced gaps in support between Donald Trump and Kamala Harris, depending on the specific poll.

@@

# Define your colors for each state (replace these colors with your desired ones)
state_colors <- c(
  alabama = dark_red,
  arizona = purple,
  arkansas = dark_red,
  california = dark_blue,
  colorado = dark_blue,
  connecticut = dark_blue,
  delaware = dark_blue,
  `district of columbia` = dark_blue,
  florida = light_blue,
  georgia = purple,
  idaho = dark_red,
  illinois = dark_blue,
  indiana = dark_red,
  iowa = light_red,
  kansas = dark_red,
  kentucky = dark_red,
  louisiana = dark_red,
  maine = "gray",
  maryland = dark_blue,
  massachusetts = dark_blue,
  michigan = purple,
  minnesota = light_blue,
  mississippi = dark_red,
  missouri = dark_red,
  montana = dark_red,
  nebraska = "gray",
  nevada = purple,
  `new hampshire` = light_blue,
  `new jersey` = dark_blue,
  `new mexico` = dark_blue,
  `new york` = dark_blue,
  `north carolina` = purple,
  `north dakota` = dark_red,
  ohio = light_red,
  oklahoma = dark_red,
  oregon = dark_blue,
  pennsylvania = purple,
  `rhode island` = dark_blue,
  `south carolina` = dark_red,
  `south dakota` = dark_red,
  tennessee = dark_red,
  texas = light_red,
  utah = dark_red,
  vermont = dark_blue,
  virginia = light_blue,
  washington = dark_blue,
  `west virginia` = dark_red,
  wisconsin = purple,
  wyoming = dark_red
)

@@

@@

ggplot(states_map, aes(x = long, y = lat, group = group, fill = region)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  scale_fill_manual(values = state_colors, 
                    breaks = plotclr,
                    labels = legend_description,
                    name = "Election Status") +  
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

```
### Legend

```{r}
#| label: tbl-electionmaplegend
#| tbl-cap: Interpretation of Electoral College Map
#| echo: false
#| warning: false

legend_tibble <- tibble(
  Color = c("Dark Red", "Light Red", "Purple", "Light Blue", "Dark Blue", "Gray"),
  Status = c("Republican", "Lean Republican", "Toss-Up", "Lean Democrat", "Democrat", "Split By District")
)

legend_tibble %>% kable(col.names = c("color", "status"))

>>>>

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  group_by(state, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "State",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

@@

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  filter(!is.na(methodology)) %>% 
  group_by(methodology, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = methodology, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "Methodology",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

```
Figure @fig-avg-pct-methodology indicates minimal bias in polling percentages when comparing different polling methodologies, as most polls show an even distribution of support between candidates. However, polls only incorporate IVR (Interactive Voice Response) tend to skew towards Harris

@@

# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

state_colors <- c(
  alabama = dark_red,
  arizona = purple,
  arkansas = dark_red,
  california = dark_blue,
  colorado = dark_blue,
  connecticut = dark_blue,
  delaware = dark_blue,
  `district of columbia` = dark_blue,
  florida = light_blue,
  georgia = purple,
  idaho = dark_red,
  illinois = dark_blue,
  indiana = dark_red,
  iowa = light_red,
  kansas = dark_red,
  kentucky = dark_red,
  louisiana = dark_red,
  maine = "gray",
  maryland = dark_blue,
  massachusetts = dark_blue,
  michigan = purple,
  minnesota = light_blue,
  mississippi = dark_red,
  missouri = dark_red,
  montana = dark_red,
  nebraska = "gray",
  nevada = purple,
  `new hampshire` = dark_blue,
  `new jersey` = dark_blue,
  `new mexico` = dark_blue,
  `new york` = dark_blue,
  `north carolina` = purple,
  `north dakota` = dark_red,
  ohio = light_red,
  oklahoma = dark_red,
  oregon = dark_blue,
  pennsylvania = purple,
  `rhode island` = dark_blue,
  `south carolina` = dark_red,
  `south dakota` = dark_red,
  tennessee = dark_red,
  texas = light_red,
  utah = dark_red,
  vermont = dark_blue,
  virginia = light_blue,
  washington = dark_blue,
  `west virginia` = dark_red,
  wisconsin = purple,
  wyoming = dark_red
)

@@

state_colors <- setNames(state_colors, tolower(names(state_colors)))

color_to_status <- setNames(names(state_colors_legend), state_colors_legend)

states_map <- states_map %>%
  mutate(status = color_to_status[state_colors[region]])

@@

ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  scale_fill_manual(
    values = state_colors_legend,
    name = "Election Status"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )
